### 2023-01-26 Strydom_et_al_2023_graph_embedding_draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@41a86090a2b482457cb251048e354194819eea58 🚀

@@

\hypertarget{graph-embeddings-produce-novel-latent-variables-to-predict-interactions}{%
\subsection{Graph embeddings produce novel latent variables to predict
interactions}\label{graph-embeddings-produce-novel-latent-variables-to-predict-interactions}}

Rather than directly predicting biological rules (see \emph{e.g.}
Pichler et al., 2020 for an overview), which may be confounded by the
sparse nature of graph data, learning embeddings works in the
low-dimensional space that maximizes information about the network
structure. This approach is further justified by the observation, for
example, that the macro-evolutionary history of a network is adequately
represented by some graph embeddings {[}Random dot product graphs
(RDPG); see Dalla Riva \& Stouffer (2016){]}. In a recent publication,
Strydom et al. (2022) have used an embedding (based on RDPG) to project
a metaweb of trophic interactions between European mammals, and
transferred this information to mammals of Canada, using the
phylogenetic distance between related clades to infer the values in the
latent sub-space into which the European metaweb was projected. By
performing the RDPG step on re-constructed values, this approach yields
a probabilistic trophic metaweb for mammals of Canada based on knowledge
of European species, despite a limited (\(\approx\) 5\%) taxonomic
overlap.

>>>>

\hypertarget{graph-embeddings-produce-latent-variables-but-not-traits}{%
\subsection{Graph embeddings produce latent variables (but not
traits)}\label{graph-embeddings-produce-latent-variables-but-not-traits}}

Rather than directly predicting biological rules (see \emph{e.g.}
Pichler et al., 2020 for an overview), which may be confounded by the
sparse nature of graph data, learning embeddings works in the
low-dimensional space that maximizes information about the network
structure. This approach is further justified by the observation, for
example, that the macro-evolutionary history of a network is adequately
represented by some graph embeddings {[}Random dot product graphs
(RDPG); see Dalla Riva \& Stouffer (2016){]}. In a recent publication,
Strydom et al. (2022) have used an embedding (based on RDPG) to project
a metaweb of trophic interactions between European mammals, and
transferred this information to mammals of Canada, using the
phylogenetic distance between related clades to infer the values in the
latent sub-space into which the European metaweb was projected. By
performing the RDPG step on re-constructed values, this approach yields
a probabilistic trophic metaweb for mammals of Canada based on knowledge
of European species, despite a limited (\(\approx\) 5\%) taxonomic
overlap, and illustrates how the values derived from an embedding can be
used for prediction without being ``traits'' of the species they
represent.

@@
